---
title: "CT"
author: "Jens Daniel Müller & Lara Burchardt"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---


```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r library, message=FALSE, warning=FALSE}

library(tidyverse)
library(ncdf4)
library(vroom) #'rlang' needs to be installed
library(lubridate)
library(geosphere)
library(dygraphs)
library(xts)
library(here)
library(metR)
library(seacarb)
#library(oce)
#library(patchwork)

```

```{r subsetting_criteria}

# route
select_route <- "E"

# variable names in 2d and 3d GETM files
var <- "SSS_east"

# latitude limits
low_lat <- 57.5
high_lat <- 58.8

```

# CT 

The mean salinity was calculated across all measurments made between march - september in the NGS subregion. 


```{r read_sss_data_fm_ngs}


nc <- nc_open(paste("data/Finnmaid/", "FM_all_2019_on_standard_tracks.nc", sep = ""))

# read required vectors from netcdf file
route <- ncvar_get(nc, "route")
route <- unlist(strsplit(route, ""))
date_time <- ncvar_get(nc, "time")
latitude_east <- ncvar_get(nc, "latitude_east")

array <- ncvar_get(nc, var) # store the data in a 2-dimensional array
#dim(array) # should have 2 dimensions: 544 coordinate, 2089 time steps
  
  fillvalue <- ncatt_get(nc, var, "_FillValue")
  array[array == fillvalue$value] <- NA
  rm(fillvalue)
  
  #i <- 5
  for (i in seq(1,length(route),1)){
  
      
    if(route[i] == select_route) {
      slice <- array[i,]
      
      value <- mean(slice[latitude_east > low_lat & latitude_east < high_lat], na.rm = TRUE)
      sd    <- sd(slice[latitude_east > low_lat & latitude_east < high_lat], na.rm = TRUE)
      date <- ymd("2000-01-01") + date_time[i]
      
      temp <- bind_cols(date = date, var=var, value = value, sd = sd)
      
      if (exists("timeseries", inherits = FALSE)){
        timeseries <- bind_rows(timeseries, temp)
      } else{timeseries <- temp}
      
      rm(temp, value, date, sd)
      
    } 
  }
nc_close(nc)

fm_sss__ngs <- timeseries %>% 
  mutate(sss = value,
         year = year(date),
         month = month(date))

fm_sss_ngs_monthlymean <- fm_sss__ngs %>% 
  filter(month >=3 , month <=9) %>% 
  summarise(sss_mean = mean(sss, na.rm = TRUE))


rm(array,fm_sss__ngs,nc, timeseries, date_time, 
  i, latitude_east, route, select_route, slice, var)
```

The mean salinity between March and September for the NGS subregion for all years is `r round(fm_sss_ngs_monthlymean,2)`.

# CT Plot

In the following section, we create a dygraph with two y-axis, combining the pCO2 measurments from VOS Finnmaid with the MLD5 calculations of the GETM Model.

```{r plot_mld__pco2_data_ngs}
gt_mld_fm_pco2_ngs <-
vroom::vroom(here::here("data/_merged_data_files/", file = "gt_mld_fm_pco2_ngs.csv"))

gt_mld_fm_pco2_ngs <- gt_mld_fm_pco2_ngs %>%
  mutate(as.POSIXct(date))
  

pco2 <- xts(cbind(gt_mld_fm_pco2_ngs$value_pCO2, gt_mld_fm_pco2_ngs$min_pCO2, gt_mld_fm_pco2_ngs$max_pCO2), order.by = gt_mld_fm_pco2_ngs$date)
names(pco2) <- c("pCO2 Finnmaid", "lwr", "upr")

mld5 <- xts(gt_mld_fm_pco2_ngs$value_mld5, order.by = gt_mld_fm_pco2_ngs$date)

plotdata <- cbind(pco2,mld5)

dygraph(plotdata) %>% 
  dySeries("mld5", axis = 'y2') %>% 
  dySeries(c("lwr","pCO2.Finnmaid", "upr")) %>% 
  dyRangeSelector(dateWindow = c("2014-01-01", "2016-12-31")) %>% 
  dyOptions(drawPoints = TRUE, pointSize = 1.5, connectSeparatedPoints=TRUE, strokeWidth=0.5,
            drawAxesAtZero=TRUE)
rm(pco2, mld5, plotdata, fm_sss_ngs_monthlymean, gt_mld_fm_pco2_ngs)
```

# Data extraction all routes

```{r fm_ngs_all_routes}
nc <- nc_open(paste("data/Finnmaid/", "FM_all_2019_on_standard_tracks.nc", sep = ""))

# read required vectors from netcdf file
route <- ncvar_get(nc, "route")
route <- unlist(strsplit(route, ""))
date_time <- ncvar_get(nc, "time")
latitude_east <- ncvar_get(nc, "latitude_east")

for (var in c("SST_east", "SSS_east", "pCO2_east")) {
  
  #print(var)
  
  array <- ncvar_get(nc, var) # store the data in a 2-dimensional array
  #dim(array) # should have 2 dimensions: 544 coordinate, 2089 time steps
  
  fillvalue <- ncatt_get(nc, var, "_FillValue")
  array[array == fillvalue$value] <- NA
  rm(fillvalue)
  
    for (i in seq(1,length(route),1)){
  
      
    if(route[i] != "N" & route[i] != "P") {
      slice <- array[i,]
      
      value <- mean(slice[latitude_east > low_lat & latitude_east < high_lat], na.rm = TRUE)
      sd    <- sd(slice[latitude_east > low_lat & latitude_east < high_lat], na.rm = TRUE)
      date <- ymd("2000-01-01") + date_time[i]
      
      fm_ngs_all_routes_part <- bind_cols(date = date, var=var, value = value, sd = sd, route=route[i])
      
      if (exists("fm_ngs_all_routes", inherits = FALSE)){
        fm_ngs_all_routes <- bind_rows(fm_ngs_all_routes, fm_ngs_all_routes_part)
      } else{fm_ngs_all_routes <- fm_ngs_all_routes_part}
      
      rm(fm_ngs_all_routes_part, value, date, sd, slice)
      
      } 
    }
  rm(array, var,i)
}   
      
nc_close(nc)

fm_ngs_all_routes %>%  
  vroom_write(here::here("data/_summarized_data_files/", file = "fm_ngs_all_routes.csv"))

rm(nc, fm_ngs_all_routes, latitude_east, route,date_time)

```

#Windspeed


```{r file_list_GETM_CT, eval=FALSE}

filesList_2d <- list.files(path= "data", pattern = "Finnmaid.E.2d.20", recursive = TRUE)

```


```{r read_getm_2d_mld_wind, eval=FALSE}
var_all <- c("u10", "v10")

filesList_2d <- list.files(path= "data", pattern = "Finnmaid.E.2d.20", recursive = TRUE)

file <- filesList_2d[1]
nc <- nc_open(paste("data/", file, sep = ""))

lon <- ncvar_get(nc, "lonc")
lat <- ncvar_get(nc, "latc", verbose = F)
nc_close(nc)

rm(file, nc)

for (a in 1:2){

  var <- var_all[a]
  
for (n in 1:length(filesList_2d)) {

file <- filesList_2d[n]

nc <- nc_open(paste("data/", file, sep = ""))

time_units <- nc$dim$time$units %>%     #we read the time unit from the netcdf file to calibrate the time 
    substr(start = 15, stop = 33) %>%   #calculation, we take the relevant information from the string
    ymd_hms()                           # and transform it to the right format

t <- time_units + ncvar_get(nc, "time")

array <- ncvar_get(nc, var) # store the data in a 2-dimensional array
dim(array) # should be 2d with dimensions: 544 coordinate, 31d*(24h/d/3h)=248 time steps

array <- as.data.frame(t(array), xy=TRUE)
array <- as_tibble(array)

gt_windspeed_ngs_part <- array %>%
  set_names(as.character(lat)) %>%
  mutate(date_time = t) %>%
  gather("lat", "value", 1:length(lat)) %>%
  mutate(lat = as.numeric(lat)) %>%
  filter(lat > low_lat, lat<high_lat) %>%
  group_by(date_time) %>%
  summarise_all("mean") %>%
  ungroup() %>%
  mutate(var = var)


if (exists("gt_windspeed_ngs")) {gt_windspeed_ngs <- bind_rows(gt_windspeed_ngs, gt_windspeed_ngs_part)
  }else {gt_windspeed_ngs <- gt_windspeed_ngs_part}

  nc_close(nc)
  rm(array, nc, t, gt_windspeed_ngs_part)
  print(n) # to see working progress
}

print(paste("gt_", var, "_ngs.csv", sep = "")) # to see working progress


rm( n, file, time_units, var)
}

rm(a, filesList_2d, var_all, lat, lon)


gt_windspeed_ngs <- gt_windspeed_ngs %>% 
  group_by(date_time, lat, var) %>% 
  summarise(mean_value= mean(value)) %>% 
  pivot_wider(values_from = mean_value, names_from = var) %>% 
  mutate(U_10 = round(sqrt(u10^2 + v10^2), 3)) %>% 
  select(-c(u10, v10))

gt_windspeed_ngs %>% 
  vroom_write(here::here("data/_summarized_data_files/", file = paste("gt_windspeed_ngs.csv", sep = "")))

```


# C~T~ calculation

C~T~ was calculated from measured pCO~2~ based on a fixed mean alkalinity value of 1650 µmol kg-1.

```{r calculate_CT_from_pCO2_Finnmaid, eval=FALSE}

df <- vroom::vroom(here::here("data/_summarized_data_files/", file = "fm_ngs_all_routes.csv"))

df <- df %>% 
  select(date, var, value, route) 

df <- df %>%
 pivot_wider(values_from = value, names_from = var) %>% 
  drop_na()

#df <- df%>%
#  mutate(sal = SSS_east,
#         tem = SST_east,
#         pCO2 = pCO2_east) %>%
#  select(pCO2,tem, sal, date)

# not enough RAM for pivot_wider
# alternative with "filter"

# df_sst <- df %>% 
#   filter( var == "SST_east") %>% 
#   mutate( tem = value) %>% 
#   select (tem)
# 
# df_sss <- df %>% 
#   filter( var == "SSS_east") %>% 
#   mutate( sal = value) %>% 
#   select( sal)
# 
# df_pco2 <- df %>% 
#   filter( var == "pCO2_east") %>% 
#   mutate (pCO2 = value) %>% 
#   select( pCO2, date)
# 
# df <- cbind(df_pco2, df_sst, df_sss) %>% 
#   select(date, tem, sal, pCO2)

# df %>% 
#   vroom_write(here::here("data/_merged_data_files/", file = paste("fm_gt__sal_tem_pco2_ngs.csv", sep = "")))

df <- df %>%
  rename(SST = SST_east, SSS = SSS_east, pCO2 = pCO2_east) %>% 
  mutate(CT = carb(24, var1=pCO2, var2=1650*1e-6,
                   S=SSS, T=SST, k1k2="m10", kf="dg", ks="d",
                   gas="insitu")[,16]*1e6)


df %>% 
   vroom_write(here::here("data/_summarized_data_files/", file = "fm_CT_ngs.csv"))

rm(df, df_sss, df_sst, df_pco2)

```

# Air-sea CO~2~ flux

The CO~2~ flux across the sea surface was calculated according to Wanninkhof (2014).


```{r calculate_gas_exchange_from_Finnmaid_pCO2_S_T_and_GETM_wind, eval=FALSE}

df_1 <- vroom::vroom(here::here("data/_summarized_data_files/", file = "fm_CT_ngs.csv"))
df_2 <- vroom::vroom(here::here("data/_summarized_data_files/", file = "gt_windspeed_ngs.csv"))

df_2$date_time <- round_date(df_2$date_time, unit = "day")
df_2 <- df_2 %>% 
  mutate(date = as.Date(date_time)) %>% 
  select(date, U_10) %>% 
  group_by(date) %>% 
  summarise_all("mean")


df <- inner_join(df_1, df_2, by = "date")
rm(df_1,df_2)

df <- df %>% 
  mutate(year = year(date))

Sc_W14 <- function(tem) {
  2116.8 - 136.25 * tem + 4.7353 * tem^2 - 0.092307 * tem^3 + 0.0007555 * tem^4
}

Sc_W14(20)

# calculate flux F [mol m–2 d–1]
df <- df %>%
  mutate(pCO2_air = 400 - 2*(2015-year),
         dpCO2 = pCO2 - pCO2_air,
         dCO2  = dpCO2 * K0(S=6.92, T=SST),
         k     = 0.251 * U_10^2 * (Sc_W14(SST)/660)^(-0.5),
         flux_daily = k*dCO2*1e-5*24)


df %>% 
   vroom_write(here::here("data/_merged_data_files/", file = paste("gt_fm_flux_ngs.csv", sep = "")))

rm(df, Sc_W14)
```

# Visualization of CT, CO2 flux and windspeeds

```{r visualization_CT_flux_windspeeds, fig.asp=0.5}

gt_fm_flux_ngs <- vroom::vroom(here::here("data/_merged_data_files/", file = "gt_fm_flux_ngs.csv"))
fm_CT_ngs <- vroom::vroom(here::here("data/_summarized_data_files/", file = "fm_CT_ngs.csv"))
gt_windspeed_ngs <- vroom::vroom(here::here("data/_summarized_data_files/", file = "gt_windspeed_ngs.csv"))


ts_xts_flux <- xts(gt_fm_flux_ngs$flux_daily, order.by = gt_fm_flux_ngs$date)
names(ts_xts_flux) <- "Daily Flux"
ts_xts_CT <- xts(fm_CT_ngs$CT, order.by = fm_CT_ngs$date)
names(ts_xts_CT) <- "CT"
ts_xts_windspeed <- xts(gt_windspeed_ngs$U_10, order.by = gt_windspeed_ngs$date_time)
names(ts_xts_windspeed) <- "Windspeed"

ts_xts_flux %>% 
  dygraph(group = "Fluxes") %>% 
  dyRangeSelector(dateWindow = c("2014-01-01", "2016-12-31")) %>% 
  dySeries("Daily Flux") %>% 
  dyAxis("y", label = "Daily Flux") %>% 
  dyOptions(drawPoints = TRUE, pointSize = 1.5, connectSeparatedPoints=TRUE, strokeWidth=0.5)

ts_xts_CT %>% 
  dygraph(group = "Fluxes") %>% 
  dyRangeSelector(dateWindow = c("2014-01-01", "2016-12-31")) %>% 
  dySeries("CT") %>% 
  dyAxis("y", label = "CT") %>% 
  dyOptions(drawPoints = TRUE, pointSize = 1.5, connectSeparatedPoints=TRUE, strokeWidth=0.5,
            drawAxesAtZero=TRUE)

ts_xts_windspeed %>% 
  dygraph(group = "Fluxes") %>% 
  dyRangeSelector(dateWindow = c("2014-01-01", "2016-12-31")) %>% 
  dySeries("Windspeed") %>% 
  dyAxis("y", label = "Windspeed [m/s]") %>% 
  dyOptions(drawPoints = TRUE, pointSize = 1.5, connectSeparatedPoints=TRUE, strokeWidth=0.5,
            drawAxesAtZero=TRUE)


rm(gt_fm_flux_ngs, gt_windspeed_ngs, fm_CT_ngs, ts_xts_CT, ts_xts_flux, ts_xts_windspeed)
```
