---
title: "CT"
author: "Jens Daniel Müller & Lara Burchardt"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---


```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r library, message=FALSE, warning=FALSE}

library(tidyverse)
library(ncdf4)
library(vroom) #'rlang' needs to be installed
library(lubridate)
library(geosphere)
library(dygraphs)
library(xts)
library(here)
library(metR)
library(oce)
library(patchwork)

```

```{r subsetting_criteria}

# route
select_route <- "E"

# variable names in 2d and 3d GETM files
var <- "SSS_east"

# latitude limits
low_lat <- 57.5
high_lat <- 58.5

```

# CT 

The mean salinity was calculated across all measurments made between march - september in the NGS subregion. 

```{r file_list_GETM_CT, eval=FALSE}

filesList_2d <- list.files(path= "data", pattern = "Finnmaid.E.2d.20", recursive = TRUE)

```

```{r read_sss_data_fm_ngs}


nc <- nc_open(paste("data/Finnmaid/", "FM_all_2019_on_standard_tracks.nc", sep = ""))

# read required vectors from netcdf file
route <- ncvar_get(nc, "route")
route <- unlist(strsplit(route, ""))
date_time <- ncvar_get(nc, "time")
latitude_east <- ncvar_get(nc, "latitude_east")

array <- ncvar_get(nc, var) # store the data in a 2-dimensional array
#dim(array) # should have 2 dimensions: 544 coordinate, 2089 time steps
  
  fillvalue <- ncatt_get(nc, var, "_FillValue")
  array[array == fillvalue$value] <- NA
  rm(fillvalue)
  
  #i <- 5
  for (i in seq(1,length(route),1)){
  
      
    if(route[i] == select_route) {
      slice <- array[i,]
      
      value <- mean(slice[latitude_east > low_lat & latitude_east < high_lat], na.rm = TRUE)
      sd    <- sd(slice[latitude_east > low_lat & latitude_east < high_lat], na.rm = TRUE)
      date <- ymd("2000-01-01") + date_time[i]
      
      temp <- bind_cols(date = date, var=var, value = value, sd = sd)
      
      if (exists("timeseries", inherits = FALSE)){
        timeseries <- bind_rows(timeseries, temp)
      } else{timeseries <- temp}
      
      rm(temp, value, date, sd)
      
    } 
  }
nc_close(nc)

fm_sss__ngs <- timeseries %>% 
  mutate(sss = value,
         year = year(date),
         month = month(date))

fm_sss_ngs_monthlymean <- fm_sss__ngs %>% 
  filter(month >=3 , month <=9) %>% 
  summarise(sss_mean = mean(sss, na.rm = TRUE))


rm(array,fm_sss__ngs,nc, timeseries, date_time, filesList_2d, high_lat, i, latitude_east, low_lat, route, select_route, slice, var )
```

The mean salinity between March and September for the NGS subregion for all years is `r round(fm_sss_ngs_monthlymean,2)`.

# CT Plot

In the following section, we create a dygraph with two y-axis, combining the pCO2 measurments from VOS Finnmaid with the MLD5 calculations of the GETM Model.

```{r plot_mld__pco2_data_ngs}
gt_mld_fm_pco2_ngs <-
vroom::vroom(here::here("data/_merged_data_files/", file = "gt_mld_fm_pco2_ngs.csv"))

gt_mld_fm_pco2_ngs <- gt_mld_fm_pco2_ngs %>%
  mutate(as.POSIXct(date))
  

pco2 <- xts(cbind(gt_mld_fm_pco2_ngs$value_pCO2, gt_mld_fm_pco2_ngs$min_pCO2, gt_mld_fm_pco2_ngs$max_pCO2), order.by = gt_mld_fm_pco2_ngs$date)
names(pco2) <- c("pCO2 Finnmaid", "lwr", "upr")

mld5 <- xts(gt_mld_fm_pco2_ngs$value_mld5, order.by = gt_mld_fm_pco2_ngs$date)

plotdata <- cbind(pco2,mld5)

dygraph(plotdata) %>% 
  dySeries("mld5", axis = 'y2') %>% 
  dySeries(c("lwr","pCO2.Finnmaid", "upr")) %>% 
  dyRangeSelector(dateWindow = c("2014-01-01", "2016-12-31")) %>% 
  dyOptions(drawPoints = TRUE, pointSize = 1.5, connectSeparatedPoints=TRUE, strokeWidth=0.5,
            drawAxesAtZero=TRUE)
rm(pco2, mld5, plotdata, fm_sss_ngs_monthlymean, gt_mld_fm_pco2_ngs)
```

# Data extraction all routes

```{r fm_ngs_all_routes}
nc <- nc_open(paste("data/Finnmaid/", "FM_all_2019_on_standard_tracks.nc", sep = ""))

# read required vectors from netcdf file
route <- ncvar_get(nc, "route")
route <- unlist(strsplit(route, ""))
date_time <- ncvar_get(nc, "time")
latitude_east <- ncvar_get(nc, "latitude_east")

for (var in c("SST_east", "SSS_east", "pCO2_east")) {
  
  print(var)
  
  array <- ncvar_get(nc, var) # store the data in a 2-dimensional array
  #dim(array) # should have 2 dimensions: 544 coordinate, 2089 time steps
  
  fillvalue <- ncatt_get(nc, var, "_FillValue")
  array[array == fillvalue$value] <- NA
  rm(fillvalue)
  
    for (i in seq(1,length(route),1)){
  
      
    if(route[i] != "N" & route[i] != "P") {
      slice <- array[i,]
      
      value <- mean(slice[latitude_east > low_lat & latitude_east < high_lat], na.rm = TRUE)
      sd    <- sd(slice[latitude_east > low_lat & latitude_east < high_lat], na.rm = TRUE)
      date <- ymd("2000-01-01") + date_time[i]
      
      fm_ngs_all_routes_part <- bind_cols(date = date, var=var, value = value, sd = sd, route=route[i])
      
      if (exists("fm_ngs_all_routes", inherits = FALSE)){
        fm_ngs_all_routes <- bind_rows(fm_ngs_all_routes, fm_ngs_all_routes_part)
      } else{fm_ngs_all_routes <- fm_ngs_all_routes_part}
      
      rm(fm_ngs_all_routes_part, value, date, sd, slice)
      
      } 
    }
  rm(array, var,i)
}   
      
nc_close(nc)

fm_ngs_all_routes %>%  
  vroom_write(here::here("data/_summarized_data_files/", file = "fm_ngs_all_routes.csv"))

rm(nc, fm_ngs_all_routes, latitude_east, route,date_time)

```
#Windspeed

```{r read_getm_2d_mld_wind, eval=FALSE}
nc_2d <- nc_open(here("data/GETM", "Finnmaid.E.2d.2018.nc"))
#print(nc_2d)
lat <- ncvar_get(nc_2d, "latc")
time_units <- nc_2d$dim$time$units %>%     #we read the time unit from the netcdf file to calibrate the time 
    substr(start = 15, stop = 33) %>%   #calculation, we take the relevant information from the string
    ymd_hms()                           # and transform it to the right format
t <- time_units + ncvar_get(nc_2d, "time") # read time vector
rm(time_units)
for (var in c("mld_rho", "u10", "v10")) {
  
#var <- "mld_rho"
array <- ncvar_get(nc_2d, var) # store the data in a 3-dimensional array
fillvalue <- ncatt_get(nc_2d, var, "_FillValue")
array[array == fillvalue$value] <- NA
array <- as.data.frame(t(array), xy=TRUE)
array <- as_tibble(array)
      
  gt_2d_part <- array %>%
  set_names(as.character(lat)) %>%
  mutate(date_time = t) %>%
  gather("lat", "value", 1:length(lat)) %>%
  mutate(lat = as.numeric(lat)) %>%
  filter(lat > low_lat, lat<high_lat) %>%
  select(-lat) %>% 
  group_by(date_time) %>% 
  summarise_all(list(value=~mean(.,na.rm=TRUE))) %>%
  ungroup() %>% 
  mutate(var = var)
     
  if (exists("gt_2d")) {
    gt_2d <- bind_rows(gt_2d, gt_2d_part)
    } else {gt_2d <- gt_2d_part} 
rm(array, fillvalue, gt_2d_part)
}
nc_close(nc_2d)
rm(nc_2d)
gt_2d <- gt_2d %>% 
  pivot_wider(values_from = value, names_from = var) %>% #doesn't know pivot_wider...
  mutate(U_10 = (sqrt(u10^2 + v10^2))) %>% 
  select(-c(u10, v10))
gt_2d %>% 
  vroom_write((here::here("data/_summarized_data_files", file = "gt_2d_windspeed.csv")))
rm(t, var, gt_2d, lat)
```
