---
title: "Finnmaid-GETM"
author: "Jens Daniel Müller & Lara Burchardt"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```


# Aim of the project

Two datasets are compared describing the sea surface temperature (SST) in the Central Baltic Sea. Observational data are provided by the VOS Finnmaid, measuring (amongst others) SST, pCO~2~, O~2~, and sea surface salinity (SSS) every minute while commuting between Travemünde and Helsinki. Measurments are taken from the surface water (3m).

The second dataset comprises GETM model data providing information and values for all water depths and locations along the Finnmaid track. 

Data are saved in the netcdf format. We describe how to read in the data, compare the relevant variables to finally visualize the relation between both SST estimates.

The documentation of the project is realized with `workflowr`, combining `R Markdown` with the version control Github as input for this webside. The following packages are needed.

```{r library, message=FALSE, warning=FALSE} 
library(tidyverse)
library(ncdf4)
library(lubridate)
library(geosphere)
```


# Definition of subsetting criteria

```{r subsetting_criteria}

select_route <- "E"
low_lat <- 57.5
high_lat <- 58.5
var <- "SST"

```

This analysis is performed for observation along the Finnmaid route `r select_route`, and in a first approach restricted to the daily mean values in the latitude range `r low_lat` - `r high_lat` deg N. The investigated variable is `r var`.


# Prepare GETM data

As a first step we want to prepare the data from the GETM model to compare it later.  We have some specifications, which files we want to look at first, because we only look at the models from a specific track, which is the Finnmaid track, route E. We want a list of all the files in our modeldata set, that include the `pattern = "Finnmaid.E.2d"`. This pattern is to be searched for in all folders and subfolders of the current working directory, we achieve that by setting `recursive = TRUE`. 

```{r file_list_GETM}

filesList <- list.files(path= "data", pattern = "Finnmaid.E.2d", recursive = TRUE)

```


We now loop through all the files in `fileList` we created before, to perform the data preparations and save the new arrays.
To open a file in the netcdf format we use the openning function of the `ncdf4` package.

```{r read_GETM_data}

for (n in 7:length(filesList)) {

file <- filesList[8]
file <- filesList[n]

nc <- nc_open(paste("data/", file, sep = ""))

#lon <- ncvar_get(nc, "lonc")
lat <- ncvar_get(nc, "latc", verbose = F)

time_units <- nc$dim$time$units %>%     #we read the time unit from the netcdf file to calibrate the time 
    substr(start = 15, stop = 33) %>%   #calculation, we take the relevant information from the string
    ymd_hms()                           # and transform it to the right format

#t <- ymd_hms("2006-1-1 00:00:00") + ncvar_get(nc, "time")
t <- time_units + ncvar_get(nc, "time")

array <- ncvar_get(nc, var) # store the data in a 3-dimensional array
dim(array) # should be 2d with dimensions: 1575 coordinate, 31d*(24h/d/3h)=248 time steps

array <- as.data.frame(t(array), xy=TRUE)
array <- as_tibble(array)


SST_GETM <- array %>%
  set_names(as.character(lat)) %>%
  mutate(date_time = t) %>%
  gather("lat", "value", 1:length(lat)) %>%
  mutate(lat = as.numeric(lat)) %>%
  filter(lat > low_lat, lat<high_lat) %>%
  group_by(date_time) %>%
  summarise_all("mean") %>%
  ungroup() %>%
  mutate(var = var)


if (exists("SST_GETM_whole")) {SST_GETM_whole <- bind_rows(SST_GETM_whole, SST_GETM)}
        else {SST_GETM_whole <- SST_GETM}

nc_close(nc)
rm(array, nc, t, lat, lon, SST_GETM)

}

# using tidyverse write_csv
# write only one data file
# use relative path with here package

SST_GETM_whole %>% 
  write_csv(here::here("data/_merged_data_files/", file = "GETM_SST_whole.csv"))

rm(SST_GETM_whole, n, file, filesList, time_units)

```



# Prepare Finnmaid data

We want to compare the model data to the data actually measured by the VOS Finnmaid. Therefore, we need to prepare the Finnmaid data next.
The file we open is called "FM_all_2019_on_standard_tracks.nc" and contains information for the time between 2003 and 2019 for route E.
The data from the VOS Finnmaid are read in from a netcdf file created by Bittig & Müller.


```{r read_finnmaid_data, echo=T, results='hide'}

#nc <- nc_open(file.choose())
nc <- nc_open(paste("data/Finnmaid/", "FM_all_2019_on_standard_tracks.nc", sep = ""))

print(nc)
attributes(nc$var)
attributes(nc$dim)

# read required vectors from netcdf file
route <- ncvar_get(nc, "route")
route <- unlist(strsplit(route, ""))
date_time <- ncvar_get(nc, "time")
latitude_east <- ncvar_get(nc, "latitude_east")

var <- "SST_east"

  array <- ncvar_get(nc, var) # store the data in a 2-dimensional array
  #dim(array) # should have 2 dimensions: 544 coordinate, 2089 time steps
  
  fillvalue <- ncatt_get(nc, var, "_FillValue")
  array[array == fillvalue$value] <- NA
  rm(fillvalue)
  
  #i <- 5
  for (i in seq(1,length(route),1)){
    
    if(route[i] == select_route) {
      slice <- array[i,]
      
      value <- mean(slice[latitude_east > low_lat & latitude_east < high_lat], na.rm = TRUE)
      date <- ymd("2000-01-01") + date_time[i]
      
      #timeseries <- bind_cols(date = date, var=var, value = value)
      temp <- bind_cols(date = date, var=var, value = value)
      
      if (exists("timeseries", inherits = FALSE)){
        timeseries <- bind_rows(timeseries, temp)
      } else{timeseries <- temp}
      
      #rm(temp, value, date)
      
    } 
  }

nc_close(nc)
#rm(list=setdiff(ls(), c("SST_GETM", "low_lat", "high_lat")))

timeseries_mean <- timeseries %>% 
  mutate(date_time = as.POSIXct(date)) %>% 
#mean of transects from finnmaid (ID)
  group_by(date_time) %>% 
  summarise_all(list(~mean(.,na.rm = TRUE),
                     ~min(., na.rm= TRUE),
                     ~max(., na.rm= TRUE),
                     ~sd(., na.rm = TRUE))) 

#rm(array, temp, high_lat,low_lat, nc, slice, var, value)
```


# Compare SST readings

In the following we compare Finnmaid and GETM data. Values are to be compared per day. Therefore, we need to calculate means per day before we continue.


```{r calculate_daily_mean}

SST_GETM_whole <- {}
SST_GETM_mean <- {}

SST_GETM_whole <- read_csv(here::here("data/_merged_data_files/", file = "GETM_SST_whole.csv"))

#SST_GETM <- read.csv("Finnmaid_E_2d_157.csv") #file nr 157 is the "GETM-test" dataset where data for a whole year was pooled
#SST_GETM$date_time <- as.POSIXct(SST_GETM$date_time)

SST_GETM_whole$date_time <- as.POSIXct(SST_GETM_whole$date_time)
SST_GETM_whole$date_time <- round_date(SST_GETM_whole$date_time, unit = "day")

#timeseries <- timeseries %>%
#  mutate(date_time = as.POSIXct(date))


SST_GETM_mean <- SST_GETM_whole %>% 
  group_by(date_time) %>% 
  summarise_all(list(~mean(.,na.rm = TRUE)))

SST_GETM_mean$var <- "SST"


#comparison <- cbind(timeseries_mean,SST_GETM_mean)

```

Now we want to plot the SST against time for both timeseries.

```{r timeseries_plot}

#if (exists("timeseries", inherits = FALSE)){
#        timeseries <- bind_rows(timeseries, temp)
#} else{timeseries <- SST_GETM_mean}

timeseries_mean <- timeseries_mean %>% 
  mutate(date_time = as.POSIXct(date_time))

timeseries_mean %>% 
  filter(date_time >= ymd_h("2003-01-01 T00"),
         date_time <= ymd_h("2018-12-31 T24")) %>%
  ggplot()+
  geom_line(aes(date_time, value_mean, col="Finnmaid"))+
  geom_point(aes(date_time, value_mean, col="Finnmaid"))+
  geom_line(data=SST_GETM_mean, aes(date_time, value, col="GETM"))+
  scale_color_brewer(palette = "Set1", name="")+
  labs(x="Date", y="SST (deg C)", title = paste("Route ", select_route," |",low_lat,"-",high_lat,"deg N"))+
  #scale_x_datetime(date_breaks = "1 months", date_labels = "%b")+
  theme_bw()

```


# Open tasks

* Check why first 6 GETM nc files do not open

