---
title: "MLD_pCO2"
author: "Jens Daniel Müller & Lara Burchardt"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---


```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r library, message=FALSE, warning=FALSE, echo= FALSE}

library(tidyverse)
library(ncdf4)
library(vroom)
library(lubridate)
library(geosphere)
library(dygraphs)
library(xts)
library(here)
library(metR)
library(vroom)

```

```{r subsetting_criteria, echo=FALSE}

# route
select_route <- "E"

# latitude limits
low_lat <- 57.5
high_lat <- 58.5

```
# Regional pCO2 and mixed layer depths

Within this chapter we relate Finnmaid pCO~2~ observations and different mixed layer depths estimates from GETM within the Northern Gotland Sea area.

## GETM mixed layer depths

As a first step we read-in following mixed layer depths estimated from GETM:

* MLD Age 1: Penetration depth of a surface tracer injected into the surface after 1 days
* MLD Age 3: Penetration depth of a surface tracer injected into the surface after 3 days
* MLD Age 5: Penetration depth of a surface tracer injected into the surface after 5 days
* MLD Rho: Mixed layer depth based on a density criterion
* MLD Tke: Mixing layer depth based on the turbulent kinetic energy and the density stratification


```{r mld_readings, eval=FALSE}
var_all <- c("mld_age_1", "mld_age_3", "mld_age_5", "mld_rho", "mld_tke")

filesList_2d <- list.files(path= "data", pattern = "Finnmaid.E.2d", recursive = TRUE)

file <- filesList_2d[8]
nc <- nc_open(paste("data/", file, sep = ""))

lon <- ncvar_get(nc, "lonc")
lat <- ncvar_get(nc, "latc", verbose = F)
corvector <- c(1:544)
nc_close(nc)

cor_space <- as.data.frame(cbind(lon, lat, corvector))


rm(file, nc)

for (a in 1:5){

  var <- var_all[a]


for (n in 7:length(filesList_2d)) {

#file <- filesList_2d[8]
file <- filesList_2d[n]

nc <- nc_open(paste("data/", file, sep = ""))

time_units <- nc$dim$time$units %>%     #we read the time unit from the netcdf file to calibrate the time 
    substr(start = 15, stop = 33) %>%   #calculation, we take the relevant information from the string
    ymd_hms()                           # and transform it to the right format

t <- time_units + ncvar_get(nc, "time")

array <- ncvar_get(nc, var) # store the data in a 2-dimensional array
dim(array) # should be 2d with dimensions: 544 coordinate, 31d*(24h/d/3h)=248 time steps

array <- as.data.frame(t(array), xy=TRUE)
array <- as_tibble(array)

#use corvector (1:544)
mld_GETM_corvector <- array %>%
  set_names(as.character(corvector)) %>%
  mutate(date_time = t) %>%
  gather("corvector", "value", 1:length(corvector)) %>%
  mutate(corvector = as.numeric(corvector),
  date = as.Date(date_time)) %>%
  select(-date_time) %>% 
  group_by(date, corvector) %>%
  summarise_all("mean") %>%
  ungroup() %>% 
  rename(mld = value)


if (exists("mld_GETM_whole")) {mld_GETM_whole <- bind_rows(mld_GETM_whole, mld_GETM_corvector)
  }else {mld_GETM_whole <- mld_GETM_corvector}

  nc_close(nc)
  rm(array, nc, t, mld_GETM_corvector)
  print(n) # to see working progress
}

print(paste("GETM_", var, "_whole.csv", sep = "")) # to see working progress

mld_GETM_whole %>% 
  vroom_write(here::here("data/_summarized_data_files/", file = paste("GETM_", var, "_whole_corvector.csv", sep = "")))

rm(mld_GETM_whole, n, file, time_units, var)
}

rm(a, filesList_2d, var_all, lat, lon)
```


## Finnmaid pCO~2~

In the following we extract Finnmaid pCO~2~ data for the Northern Gotland Sea. For each crossing of the ferry, mean, minimum, maximum and the standard deviation of pCO~2~ are calculated.


```{r finnmaid_pCO2, eval = FALSE}
nc <- nc_open(paste("data/Finnmaid/", "FM_all_2019_on_standard_tracks.nc", sep = ""))

var <- "pCO2_east"

# read required vectors from netcdf file
route <- ncvar_get(nc, "route")
route <- unlist(strsplit(route, ""))
date_time <- ncvar_get(nc, "time")
latitude_east <- ncvar_get(nc, "latitude_east")

array <- ncvar_get(nc, var) # store the data in a 2-dimensional array
#dim(array) # should have 2 dimensions: 544 coordinate, 2089 time steps
  
  fillvalue <- ncatt_get(nc, var, "_FillValue")
  array[array == fillvalue$value] <- NA
  rm(fillvalue)
  
  for (i in seq(1,length(route),1)){
  
      
    if(route[i] == select_route) {
      slice <- array[i,]
      
      value <- mean(slice[latitude_east > low_lat & latitude_east < high_lat], na.rm = TRUE)
      sd    <- sd(slice[latitude_east > low_lat & latitude_east < high_lat], na.rm = TRUE)
      min   <- min(slice[latitude_east > low_lat & latitude_east < high_lat], na.rm = TRUE)
      max   <- max(slice[latitude_east > low_lat & latitude_east < high_lat], na.rm = TRUE)
      date <- ymd("2000-01-01") + date_time[i]
      
      temp <- bind_cols(date = date, var=var, value = value, sd = sd, min = min, max = max)
      
      if (exists("timeseries", inherits = FALSE)){
        timeseries <- bind_rows(timeseries, temp)
      } else{timeseries <- temp}
      
      rm(temp, value, date, sd, min, max)
      
    } 
  }
nc_close(nc)

timeseries$date_time <- as.POSIXct(timeseries$date) %>% 
  cut.POSIXt(breaks = "days") %>% 
  round.POSIXt(units = "days") %>% 
  as.POSIXct(tz = "UTC")
timeseries <- timeseries %>% 
  select(-c(date))

timeseries %>% 
  vroom_write(here::here("data/_summarized_data_files/", file = "Finnmaid_pCO2_whole.csv"))

rm(array, high_lat,low_lat, nc, slice, var, timeseries, latitude_east, date_time, route,i)
```
## Data merging

We want a datafile including the pCO~2~ values, as measured from the Finnmaid, as well as all five mixed layer depth parameters.

```{r merging_mld_pco2, eval=FALSE}
start_time <- Sys.time()

ts_pCO2 <- vroom::vroom(here::here("data/_summarized_data_files/", file = "Finnmaid_pCO2_whole.csv"))#, na = c("NA", "NaN", "Inf", "-Inf"))
ts_mld_age_1 <- vroom::vroom(here::here("data/_summarized_data_files/", file = "GETM_mld_age_1_whole_corvector.csv"))
ts_mld_age_3 <- vroom::vroom(here::here("data/_summarized_data_files/", file = "GETM_mld_age_3_whole_corvector.csv"))
ts_mld_age_5 <- vroom::vroom(here::here("data/_summarized_data_files/", file = "GETM_mld_age_5_whole_corvector.csv"))
ts_mld_rho <- vroom::vroom(here::here("data/_summarized_data_files/", file = "GETM_mld_rho_whole_corvector.csv"))
ts_mld_tke <- vroom::vroom(here::here("data/_summarized_data_files/", file = "GETM_mld_tke_whole_corvector.csv"))

#pCO2 Finnmaid
ts_pCO2 <- ts_pCO2 %>% 
  transmute( value_pCO2 = value, sd_pCO2 = sd, min_pCO2 = min, max_pCO2 = max, date = date_time)

#cor_space, which corvector entries belong to lat restrictions
# if no lat restrictions, comment out the following 4 lines or set them such, that they don't actually restrict

filesList_2d <- list.files(path= "data", pattern = "Finnmaid.E.2d", recursive = TRUE)

file <- filesList_2d[8]
nc <- nc_open(paste("data/", file, sep = ""))
lon <- ncvar_get(nc, "lonc")
lat <- ncvar_get(nc, "latc", verbose = F)
corvector <- c(1:544)
nc_close(nc)
cor_space <- as.data.frame(cbind(lon, lat, corvector))

cor_space_restriction <- cor_space %>% 
  filter(lat > low_lat, lat<high_lat)
low_cor <- min(cor_space_restriction$corvector)
high_cor <- max(cor_space_restriction$corvector)

#mld_age_1
ts_mld_age_1 <- ts_mld_age_1 %>%
  filter(corvector > low_cor, corvector<high_cor) %>%
  group_by(date) %>% 
  summarise_all(list(value=~mean(.,na.rm=TRUE))) %>%
  ungroup() %>%
  transmute( date = date, corvector = corvector_value, value_mld1 = mld_value)
#mld_age_3
ts_mld_age_3 <- ts_mld_age_3 %>%
  filter(corvector > low_cor, corvector<high_cor) %>%
  group_by(date) %>% 
  summarise_all(list(value=~mean(.,na.rm=TRUE))) %>%
  ungroup() %>%
  transmute( date = date, corvector = corvector_value, value_mld3 = mld_value)
#mld_age_5
ts_mld_age_5 <- ts_mld_age_5 %>%
  filter(corvector > low_cor, corvector<high_cor) %>%
  group_by(date) %>% 
  summarise_all(list(value=~mean(.,na.rm=TRUE))) %>%
  ungroup() %>%
  transmute( date = date, corvector = corvector_value, value_mld5 = mld_value)
#mld_rho
ts_mld_rho <- ts_mld_rho %>%
  filter(corvector > low_cor, corvector<high_cor) %>%
  group_by(date) %>% 
  summarise_all(list(value=~mean(.,na.rm=TRUE))) %>%
  ungroup() %>%
  transmute( date = date, corvector = corvector_value, value_mldrho = mld_value)
#mld_tke
ts_mld_tke <- ts_mld_tke %>%
  filter(corvector > low_cor, corvector<high_cor) %>%
  group_by(date) %>% 
  summarise_all(list(value=~mean(.,na.rm=TRUE))) %>%
  ungroup() %>%
  transmute( date = date, corvector = corvector_value, value_mldtke = mld_value)


ts_mld <- bind_cols(ts_mld_age_1, ts_mld_age_3, ts_mld_age_5, ts_mld_rho, ts_mld_tke)
ts_mld <- ts_mld %>% 
  select(date, corvector, value_mld1, value_mld3, value_mld5, value_mldrho, value_mldtke) %>% 
  mutate(date = as.POSIXct(date)) 

ts_mld$date <- ts_mld$date %>% 
  cut.POSIXt(breaks = "days") %>% 
  round.POSIXt(units = "days") %>% 
  as.POSIXct(tz = "UTC")

ts_mld_pco2_full <- full_join(ts_pCO2, ts_mld, by = "date")

ts_mld_pco2_full %>% 
  vroom_write(here::here("data/_merged_data_files/", file = "Finnmaid_GETM_pCO2_mld_latrestrictions.csv"))
```

## Timeseries plots

```{r timeseries_pCO2_dygraphs, fig.asp=0.5}

ts_mld_pco2 <- vroom::vroom(here::here("data/_merged_data_files/", file = "Finnmaid_GETM_pCO2_mld_latrestrictions.csv"))

ts_xts <- xts(cbind(ts_mld_pco2$value_mld1,ts_mld_pco2$value_mld3, ts_mld_pco2$value_mld5, ts_mld_pco2$value_mldrho, ts_mld_pco2$value_mldtke), order.by = ts_mld_pco2$date)
names(ts_xts) <- c("MLD Age 1", "MLD Age 3", "MLD Age 5", "MLD Rho", "MLD Tke")
ts_pco2_xts <- xts(cbind(ts_mld_pco2$value_pCO2, ts_mld_pco2$min_pCO2, ts_mld_pco2$max_pCO2), order.by = ts_mld_pco2$date)
names(ts_pco2_xts) <- c("pCO2 Finnmaid", "lwr", "upr")

ts_xts %>% 
  dygraph(group = "MLD") %>% 
  dyRangeSelector(dateWindow = c("2014-01-01", "2016-12-31")) %>% 
  dySeries("MLD Age 1") %>% 
  dySeries("MLD Age 3", color = "red") %>% 
  dyAxis("y", label = "MLD [m]") %>% 
  dyOptions(drawPoints = TRUE, pointSize = 1.5, connectSeparatedPoints=TRUE, strokeWidth=0.5)

ts_pco2_xts %>% 
  dygraph(group = "MLD") %>% 
  dyRangeSelector(dateWindow = c("2014-01-01", "2016-12-31")) %>% 
  dySeries(c("lwr", "pCO2 Finnmaid","upr")) %>% 
  dyAxis("y", label = "") %>% 
  dyOptions(drawPoints = TRUE, pointSize = 1.5, connectSeparatedPoints=TRUE, strokeWidth=0.5,
            drawAxesAtZero=TRUE)

rm(ts,ts_dif_xts,ts_xts)
```

# Basin-wide pCO2 and mixed layer depth
## GETM mixed layer depths

The dataset for the regional analysis of mixed layer depths and pCO2 were written such, that already the complete information was calculated for the GETM data.

## Finnmaid pCO~2~

The pCO~2~ measurments are also to be analysed for the whole basin next. Same as for the SST analysis, the route is divided in 2 km grids (startpoint Travemuende). That results in 544 coordinate points in every transect. The corresponding coordinate number (1:544) is added to the dataset, with the exact longitude and lattitude coordinates. 

```{r finnmaid_basinwide_pCO2, eval = FALSE}
#Prepare data for Hovmoeller Plot

nc <- nc_open(paste("data/Finnmaid/", "FM_all_2019_on_standard_tracks.nc", sep = ""))

var <- "pCO2_east"

# read required vectors from netcdf file
route <- ncvar_get(nc, "route")
route <- unlist(strsplit(route, ""))
date_time <- ncvar_get(nc, "time")
latitude_east <- ncvar_get(nc, "latitude_east")
longitude_east <-ncvar_get(nc, "longitude_east")
date_time_o <- ncvar_get(nc, "otime_east")

  array <- ncvar_get(nc, var) # store the data in a 2-dimensional array
  #dim(array) # should have 2 dimensions: 544 coordinate, 2089 time steps
  
  fillvalue <- ncatt_get(nc, var, "_FillValue")
  array[array == fillvalue$value] <- NA
  rm(fillvalue)
  
  cor_vector <- c(1:544)
  
  #i <- 5
start <- Sys.time()
  for (i in seq(1,length(route),1)){
  
      
    if(route[i] == select_route) {
      slice <- array[i,]          #define slice of the data, per row (per measurment day)
      value <- slice
      date <- ymd("2000-01-01") + date_time[i]
      
      #if detailed date/time information is needed: uncomment that
      #date <- as.Date(c(1:544))   # set up "date" variable to be overwritten later, needs to be "Date" object
      #for (a in seq(1,length(latitude_east),1)){              # for slice i the corresponding 544 transect steps are  
      #temp_time <- ymd("2000-01-01") + date_time_o[i,a]       # adjoined by corresponding time ("otime_east")
      #date[a] <- temp_time
      #}
      #temp <- bind_cols(value = value, lon = longitude_east, lat = latitude_east, 
      #corvector =   cor_vector, date_time = date)
      #temp$date_time <- as.POSIXct(temp$date_time)
      
      #when detailed time/date information is needed, comment the following
      temp <- bind_cols(value = value, lon = longitude_east, lat = latitude_east,
      corvector =  cor_vector)
      temp$date <- date
      # 
      
      if (exists("timeseries", inherits = FALSE)){
        timeseries <- bind_rows(timeseries, temp)
      } else{timeseries <- temp}
      
      rm(temp, value, date)
    }
    print(i)
  }
nc_close(nc)

timeseries <- timeseries %>% 
  mutate(value_pCO2_Finn = value) %>% 
  select(-c(value))

timeseries %>% 
  vroom_write(here::here("data/_summarized_data_files/", file = "Finnmaid_pCO2_whole_nolatlimits_corvector.csv"))

rm(array, nc, slice, var, date_time_o, cor_vector, date_time, latitude_east,longitude_east,route,i,start, timeseries)
```

## Data Merging without lat restrictions

```{r merging_mld_pco2_nolatrestriction, eval=FALSE}

ts_pCO2 <- vroom::vroom(here::here("data/_summarized_data_files/", file = "Finnmaid_pCO2_whole_nolatlimits_corvector.csv"))
ts_mld_age_1 <- vroom::vroom(here::here("data/_summarized_data_files/", file = "GETM_mld_age_1_whole_corvector.csv"))
ts_mld_age_3 <- vroom::vroom(here::here("data/_summarized_data_files/", file = "GETM_mld_age_3_whole_corvector.csv"))
ts_mld_age_5 <- vroom::vroom(here::here("data/_summarized_data_files/", file = "GETM_mld_age_5_whole_corvector.csv"))
ts_mld_rho <- vroom::vroom(here::here("data/_summarized_data_files/", file = "GETM_mld_rho_whole_corvector.csv"))
ts_mld_tke <- vroom::vroom(here::here("data/_summarized_data_files/", file = "GETM_mld_tke_whole_corvector.csv"))

ts_mld_age_1 <- ts_mld_age_1 %>%
    transmute( date = date, corvector = corvector, value_mld1 = mld)

ts_mld_age_3 <- ts_mld_age_3 %>%
    transmute( date = date, corvector = corvector, value_mld3 = mld)

ts_mld_age_5 <- ts_mld_age_5 %>%
    transmute( date = date, corvector = corvector, value_mld5 = mld)

ts_mld_rho <- ts_mld_rho %>%
    transmute( date = date, corvector = corvector, value_mldrho = mld)

ts_mld_tke <- ts_mld_tke %>%
    transmute( date = date, corvector = corvector, value_mldtke = mld)

ts_mld <- bind_cols(ts_mld_age_1, ts_mld_age_3, ts_mld_age_5, ts_mld_rho, ts_mld_tke)
ts_mld <- ts_mld %>% 
  select(date, corvector, value_mld1, value_mld3, value_mld5, value_mldrho, value_mldtke) %>% 
  mutate(date = as.POSIXct(date)) 

ts_mld$date <- ts_mld$date %>% 
  cut.POSIXt(breaks = "days") %>% 
  round.POSIXt(units = "days") %>% 
  as.POSIXct(tz = "UTC")

ts_pCO2$date <- ts_pCO2$date %>% 
  as.POSIXct() %>% 
  cut.POSIXt(breaks = "days") %>% 
  round.POSIXt(units = "days") %>% 
  as.POSIXct(tz = "UTC")

ts_mld_pco2_full <- full_join(ts_pCO2, ts_mld, by =c("date","corvector"))

ts_mld_pco2_full %>% 
vroom_write(here::here("data/_merged_data_files/", file = "Finnmaid_GETM_pCO2_mld_nolatrestrictions.csv"))
```

## Hovmoeller Plots

### Daily GETM mld

We present Hovmoeller Plots for the 5 mixed layer depth parameters (see 5.1). The daily GETM mld values are presented as a function of time and the ships distance to Travemuende along route E. Plots are shown seperatly for each parameter and years. 
This timeseries represents mixed layer depth values for the parameter `mld_age_1` from 2007 to 2019. 
```{r hovmoeller_mld_1, fig.asp=5, fig.cap="Daily mean GETM mld age 1 values as a function of time and the ships distance to Travemuende along route E."}

ts_mld_pco2_full <- 
  vroom::vroom(here::here("data/_merged_data_files/", file = "Finnmaid_GETM_pCO2_mld_nolatrestrictions.csv"), guess_max = 105537)
# why guess_max so high? because the first 105536 rows of mld values are empty and when opening are interpreted as factors, which results in a parsing error

ts_mld_pco2_full_hov <- ts_mld_pco2_full %>% 
  mutate(date = ymd(date), 
         year = year(date),
         dist.trav = corvector*2)
#mld age 1
ts_mld_pco2_full_hov %>% 
  filter(year > 2006, year < 2019) %>% 
  ggplot()+
    geom_raster(aes(date, dist.trav, fill=value_mld1))+
      scale_fill_viridis_c(name="GETM MLD Age 1 [m]",
                           limits = c(0,80),
                           direction = -1)+
      scale_x_date(expand = c(0,0))+
      scale_y_continuous(expand = c(0,0))+
      labs(y="Distance to Travemuende [km]")+
      theme_bw()+
      theme(
        axis.title.x = element_blank(),
        legend.position = "bottom",
        legend.key.width = unit(1.3, "cm"),
        legend.key.height = unit(0.3, "cm")
      )+
      facet_wrap(~year, ncol = 1, scales = "free_x")
```

This timeseries represents mixed layer depth values for the parameter `mld_age_3` from 2007 to 2019.

```{r hovmoeller_mld_3, fig.asp=5, fig.cap="Daily mean GETM mld age 3 values as a function of time and the ships distance to Travemuende along route E."}
#mld age 3
ts_mld_pco2_full_hov %>% 
  filter(year > 2006, year < 2019) %>% 
  ggplot()+
    geom_raster(aes(date, dist.trav, fill=value_mld3))+
      scale_fill_viridis_c(name="GETM MLD Age 3 [m]",
                           limits = c(0,80),
                           direction = -1)+
      scale_x_date(expand = c(0,0))+
      scale_y_continuous(expand = c(0,0))+
      labs(y="Distance to Travemuende [km]")+
      theme_bw()+
      theme(
        axis.title.x = element_blank(),
        legend.position = "bottom",
        legend.key.width = unit(1.3, "cm"),
        legend.key.height = unit(0.3, "cm")
      )+
      facet_wrap(~year, ncol = 1, scales = "free_x")
```

This timeseries represents mixed layer depth values for the parameter `mld_age_5` from 2007 to 2019.

```{r hovmoeller_mld_5, fig.asp=5, fig.cap="Daily mean GETM mld age 5 values as a function of time and the ships distance to Travemuende along route E."}
#mld age 5
ts_mld_pco2_full_hov %>% 
  filter(year > 2006, year < 2019) %>% 
  ggplot()+
    geom_raster(aes(date, dist.trav, fill=value_mld5))+
      scale_fill_viridis_c(name="GETM MLD Age 5 [m]",
                           limits = c(0,80),
                           direction = -1)+
      scale_x_date(expand = c(0,0))+
      scale_y_continuous(expand = c(0,0))+
      labs(y="Distance to Travemuende [km]")+
      theme_bw()+
      theme(
        axis.title.x = element_blank(),
        legend.position = "bottom",
        legend.key.width = unit(1.3, "cm"),
        legend.key.height = unit(0.3, "cm")
      )+
      facet_wrap(~year, ncol = 1, scales = "free_x")
```

This timeseries represents mixed layer depth values for the parameter `mld_rho` from 2007 to 2019.

```{r hovmoeller_mld_rho, fig.asp=5, fig.cap="Daily mean GETM mld rho values as a function of time and the ships distance to Travemuende along route E."}
#mld rho
ts_mld_pco2_full_hov %>% 
  filter(year > 2006, year < 2019) %>% 
  ggplot()+
    geom_raster(aes(date, dist.trav, fill=value_mldrho))+
      scale_fill_viridis_c(name="GETM MLD Rho [m]",
                           limits = c(0,80),
                           direction = -1)+
      scale_x_date(expand = c(0,0))+
      scale_y_continuous(expand = c(0,0))+
      labs(y="Distance to Travemuende [km]")+
      theme_bw()+
      theme(
        axis.title.x = element_blank(),
        legend.position = "bottom",
        legend.key.width = unit(1.3, "cm"),
        legend.key.height = unit(0.3, "cm")
      )+
      facet_wrap(~year, ncol = 1, scales = "free_x")

```

This timeseries represents mixed layer depth values for the parameter `mld_tke` from 2007 to 2019.

```{r hovmoeller_mld_tke, fig.asp=5, fig.cap="Daily mean GETM mld tke values as a function of time and the ships distance to Travemuende along route E."}
#mld tke
# there are negative values and values up to ~3000, therefore the second filter (filter(value_mldtke >0, value_mldtke < 300)) was added

ts_mld_pco2_full_hov %>% 
  filter(year > 2006, year < 2019) %>%
  #filter(value_mldtke >0, value_mldtke < 300) %>% 
  ggplot()+
    geom_raster(aes(date, dist.trav, fill=value_mldtke))+
      scale_fill_viridis_c(name="GETM MLD Tke [m]",
                           limits = c(0,80),
                           direction = -1)+
      scale_x_date(expand = c(0,0))+
      scale_y_continuous(expand = c(0,0))+
      labs(y="Distance to Travemuende [km]")+
      theme_bw()+
      theme(
        axis.title.x = element_blank(),
        legend.position = "bottom",
        legend.key.width = unit(1.3, "cm"),
        legend.key.height = unit(0.3, "cm")
      )+
      facet_wrap(~year, ncol = 1, scales = "free_x")
```

###Weekly Finnmaid pCO2 readings 
```{r hovmoeller_Finnmaid_pco2, fig.asp=5, fig.cap="Mean weekly observed (Finnmaid) pCO2 as a function of time and the ships distance to Travemuende along route E."}

ts_pco2_full_hov <- ts_mld_pco2_full %>% 
  mutate(date = ymd(date),
         week = as.Date(cut(date, breaks="weeks")),
         dist.trav = corvector*2)

ts_pco2_full_hov <- ts_pco2_full_hov %>% 
  select(value_pCO2_Finn, week, dist.trav) %>% 
  dplyr::group_by(dist.trav, week) %>% 
  summarise_all(list(mean=~mean(.,na.rm=TRUE))) %>% 
  as.data.frame() %>% 
  mutate(year = year(week),
  value_pco2_Finn_mean = mean)


ts_pco2_full_hov %>%
  filter(year > 2006, year < 2019) %>% 
    ggplot()+
    geom_raster(aes(week, dist.trav, fill=value_pco2_Finn_mean))+
    #scale_fill_scico(palette = "vik", name="mean difference in SST [°C]")+
    scale_fill_viridis_c(name="Finnmaid pCO2 [ppm]",
                         limits = c(50,600),
                         na.value = "white")+
    scale_x_date(expand = c(0,0))+
    scale_y_continuous(expand = c(0,0))+
    labs(y="Distance to Travemuende [km]")+
    theme_bw()+
    theme(
      axis.title.x = element_blank(),
      legend.position = "bottom",
      legend.key.width = unit(1.3, "cm"),
      legend.key.height = unit(0.3, "cm")
    )+
    facet_wrap(~year, ncol = 1, scales = "free_x")
```

### Comparison pCO2 and mld age 5

```{r hovmoeller_Finmmaid-GETM_pcO2-mld, fig.asp=5, fig.cap="Mean weekly difference between modelled (GETM)  mixed layer depth and observed (Finnmaid) pCO2 values as a function of time and the ships distance to Travemuende along route E."}

ts_mld_pco2_full_hov <- ts_mld_pco2_full %>% 
  mutate(date = ymd(date),
         week = as.Date(cut(date, breaks="weeks")),
         dist.trav = corvector*2)

comparison_hov_pco2_mld <- ts_mld_pco2_full_hov %>% 
  mutate(relation = (value_pCO2_Finn/value_mld5)) %>% 
  select(relation, week, dist.trav) %>% 
  dplyr::group_by(dist.trav, week) %>% 
  summarise_all(list(relation=~mean(., na.rm = TRUE))) %>% 
  as.data.frame %>% 
  mutate(year = year(week))               

comparison_hov_pco2_mld %>%
  dplyr::filter(year > 2006, year < 2019) %>% 
    ggplot()+
    geom_raster(aes(week, dist.trav, fill= relation))+
    scale_fill_viridis_c(name="pCO2 / MLD [ppm/m]",
                         limits = c(0,50),
                         na.value = "white")+
    scale_x_date(expand = c(0,0))+
    scale_y_continuous(expand = c(0,0))+
    labs(y="Distance to Travemuende [km]")+
    theme_bw()+
    theme(
      axis.title.x = element_blank(),
      legend.position = "bottom",
      legend.key.width = unit(1.3, "cm"),
      legend.key.height = unit(0.3, "cm")
    )+
    facet_wrap(~year, ncol = 1, scales = "free_x")

rm(ts_mld_pco2_full_hov, comparison_hov_pco2_mld, ts_mld_pco2_full, ts_mld_pco2_full, ts_pco2_full_hov)
```
